---
title: "Lab 11 - Grading the professor, Pt. 2"
author: "Noah Booker"
date: "4/24/25"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
```

# Part 1: Simple linear regression

## Exercise 1

> Fit a linear model (one you have fit before): m_bty, predicting average professor evaluation score based on average beauty rating (bty_avg) only. Write the linear model, and note the R-squared and the adjusted R-squared.

```{r m_bty}
m_bty <- summary(lm(score ~ bty_avg, data = evals))
m_bty
```

score = 3.89 + .07 x bty_avg + error

R-squared = .04

adjusted R-squared = .03

# Part 2: Multiple linear regression

## Exercise 2

> Fit a linear model (one you have fit before): m_bty_gen, predicting average professor evaluation score based on average beauty rating (bty_avg) and gender. Write the linear model, and note the R-squared and the adjusted R-squared.

```{r m_bty_gen}
m_bty_gen <- summary(lm(score ~ bty_avg + gender, data = evals))
m_bty_gen
```



## Additional Exercises

*Repeat the format above for additional exercises.*
